---
title: "Exploratory data analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

First of all we have to clean and format the data.

```{r}



us_customer_sample <- US_customers[c(10:21),c(1:14)]
#format dates
us_customer_sample$Date.Sold<-gsub("-","/",us_customer_sample$Date.Sold)
us_customer_sample$Date.Sold <- as.Date(us_customer_sample$Date.Sold, format = "%m/%d/%y")
#filtering by date
from_date="2014/01/01"
to_date="2014/12/01"
us_customer_sample[ (us_customer_sample$Date.Sold >= "2014/01/01" & us_customer_sample$Date.Sold <= "2014/12/01")  , ]
#alternate filteR
#filter(us_customer_sample, Date.Sold >= "2014/01/01", Date.Sold <= "2014/12/01")

```

merkes account#11640865

setup

```{r}
library(lubridate)
library(dplyr)
library(lattice)
data_frame=US_customers



#format dates
data_frame$Date.Sold<-gsub("-","/",data_frame$Date.Sold)
data_frame$Date.Sold <- as.Date(data_frame$Date.Sold, format = "%m/%d/%y")
data_frame["Year"]=year(as.Date(data_frame$Date.Sold))

```


The number of customers over the years.
```{r}
#filter out years 2014-2017
years_considered <- c("2014", "2015", "2016", "2017")
data_frame=filter(data_frame, Year %in% years_considered) 

#Number of customers over the years
table(year(as.Date(data_frame$Date.Sold)))
```

Customers in 2014
```{r}

#customers in 2014

barchart(Date.Sold~Description,data=data_frame,groups=Year,scales=list(x=list(rot=90,cex=0.8)))

```

```{r}
#how many sales have been made over the years by different contract groups

data.m <- melt(data_frame, id=c("ï..CustomerId","City","State","Dealer.Account.Number","Year", "Description"),measure.vars = c("Description"))

no_sales_by_contract_group <- cast(data.m, Year+Description~variable,length)
colnames(no_sales_by_contract_group)[3] <- "no.of.sales"



#how many sales each dealer made by year and contract group
data.m <- melt(data_frame, id=c("ï..CustomerId","City","State","Dealer.Account.Number","Year", "Description"),measure.vars = c("Dealer.Account.Number"))

data.c <- cast(data.m, Year+Description+Dealer.Account.Number~variable,length)


#how many units did each dealer deliver by year and contract group

data.m <- melt(data_frame, id=c("ï..CustomerId","City","State","Dealer.Account.Number","Year", "Description"),measure.vars = c("Units"))

data.c <- cast(data.m, Year+Description+Dealer.Account.Number~variable,sum)

g<- ggplot(data.c,aes(Year))
 g + geom_bar(aes(fill=Description), width = 0.5) + 
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
    labs(title="Categorywise Bar Chart", 
         subtitle="Manufacturer of vehicles", 
         caption="Source: Manufacturers from 'mpg' dataset")
#how many dealers have we had over the years in different contract groups

data.m <- melt(data_frame, id=c("ï..CustomerId","City","State","Dealer.Account.Number","Year", "Description"),measure.vars = c("Dealer.Account.Number"))

data.c <- cast(data.m, Year+Description+Dealer.Account.Number~variable,length)

a<-as.data.frame(data.c[,c(1,2,3)])

data.m <- melt(a, id=c("Year", "Description"),measure.vars = c("Dealer.Account.Number"))
no_of_dealers_by_contract_group<- cast(data.m, Year+Description~variable,length)
colnames(no_of_dealers_by_contract_group)[3] <- "no.of.dealers"


# pie chart of number of dealers in  year

year = 2016
par(mfrow = c(2,2))
for (year in c(2014,2015,2016,2017)){
  
  slices <-no_of_dealers_by_contract_group$no.of.dealers[no_of_dealers_by_contract_group$Year==year] 
lbls <-no_of_dealers_by_contract_group$Description[no_of_dealers_by_contract_group$Year==year] 
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices,labels = lbls, col=rainbow(length(lbls)),
  	main=paste("Dealers by contract group",year))
  
}



```

pie chart of #dealers by contract group
````{r}
par(mfrow = c(2,2))
for (year in c(2014,2015,2016,2017)){
  
  slices <-no_sales_by_contract_group$no.of.sales[no_sales_by_contract_group$Year==year] 
lbls <-no_sales_by_contract_group$Description[no_sales_by_contract_group$Year==year] 
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices,labels = lbls, col=rainbow(length(lbls)),
  	main=paste("Sales by contract group",year))
  
}

````




Pie chart of how many of the units were sold by contract group
```{r}

#first let\s find no of units by contract group

data.m <- melt(data_frame, id=c("ï..CustomerId","City","State","Dealer.Account.Number","Year", "Description"),measure.vars = c("Units"))

no_of_units_by_contract_group <- cast(data.m, Year+Description~variable,sum)
#colnames(no_of_units_by_contract_group)[3] <- "no.of.units"


par(mfrow = c(2,2))
for (year in c(2014,2015,2016,2017)){
  
  slices <-no_of_units_by_contract_group$Units[no_of_units_by_contract_group$Year==year] 
lbls <-no_of_units_by_contract_group$Description[no_of_units_by_contract_group$Year==year] 
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices,labels = lbls, col=rainbow(length(lbls)),
  	main=paste("Units by contract group",year))
}

#put everything together
units_sales_dealers_by_year<-no_of_dealers_by_contract_group
units_sales_dealers_by_year$Units<-no_of_units_by_contract_group$Units
units_sales_dealers_by_year$no.of.sales<-no_sales_by_contract_group$no.of.sales
units_sales_dealers_by_year$Average.Units<-units_sales_dealers_by_year$Units/units_sales_dealers_by_year$no.of.dealers
units_sales_dealers_by_year$Average.Package<-units_sales_dealers_by_year$Units/units_sales_dealers_by_year$no.of.sales

# replace first with 1st, second with 2nd etc.
units_sales_dealers_by_year$Description <- gsub('First', '1st', units_sales_dealers_by_year$Description)
units_sales_dealers_by_year$Description <- gsub('Second', '2nd', units_sales_dealers_by_year$Description)
units_sales_dealers_by_year$Description <- gsub('Third', '3rd', units_sales_dealers_by_year$Description)
units_sales_dealers_by_year$Description <- gsub('Fourth', '4th', units_sales_dealers_by_year$Description)
units_sales_dealers_by_year$Description <- gsub('Fifth', '5th', units_sales_dealers_by_year$Description)
units_sales_dealers_by_year$Description <- gsub('Sixth', '6th', units_sales_dealers_by_year$Description)
units_sales_dealers_by_year$Description <- gsub('Seventh', '7th', units_sales_dealers_by_year$Description)
units_sales_dealers_by_year$Description <- gsub('Eighth', '8th', units_sales_dealers_by_year$Description)
units_sales_dealers_by_year$Description <- gsub('Ninth', '9th', units_sales_dealers_by_year$Description)

```

Which of the contract groups has the highest average units
```{r}

ggplot(units_sales_dealers_by_year, aes(Year, Average.Units, fill = Description)) + 
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette = "Paired")


```
Which of the contract groups has the highest average package
```{r}

ggplot(units_sales_dealers_by_year, aes(Year, Average.Package, fill = Description)) + 
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette = "Paired")


```
Total units by years

```{r}
data.m <- melt(data_frame, id=c("ï..CustomerId","City","State","Dealer.Account.Number","Year", "Description"),measure.vars = c("Units"))

total_units_by_year <- cast(data.m, Year~variable,sum)

ggplot(total_units_by_year, aes(x=Year)) + 
  geom_line(aes(y=Units)) + 
  labs(title="Units sold by year", 
       subtitle="Returns Percentage from 'Economics' Dataset", 
       caption="Source: Southwestern Advantage", 
       y="Yearly units")


```
Let's look at production weekly by year and by contract groups
```{r}

data.m <- melt(data_frame, id=c("ï..CustomerId","City","State","Dealer.Account.Number","Year", "Description","Week.Sold"),measure.vars = c("Units"))
units_by_week <- cast(data.m, Year+Week.Sold+Description~variable,sum)
units_by_week_by_dealer <- cast(data.m, Year+Week.Sold+Description+Dealer.Account.Number~variable,sum)
sales_by_week_by_contract<-cast(data.m, Year+Week.Sold+Description~variable,length)
colnames(sales_by_week_by_contract)[4] <- "no.of.customers"

data.m<-melt(units_by_week_by_dealer, id=c("Year","Description","Week.Sold" ),measure.vars = c("Units"))
no_of_dealers_selling_weekly_by_contract= cast(data.m, Year+Week.Sold+Description~variable,length)
units_sold_by_week_by_contract_group<-cast(data.m, Year+Week.Sold+Description~variable,sum)
units_sold_by_week_by_contract_group$no.of.Dealers.selling<-no_of_dealers_selling_weekly_by_contract$Units
units_sold_by_week_by_contract_group$Average.Weekly.Units<-units_sold_by_week_by_contract_group$Units/units_sold_by_week_by_contract_group$no.of.Dealers.selling
units_sold_by_week_by_contract_group$no.of.Customers<-sales_by_week_by_contract$no.of.customers
units_sold_by_week_by_contract_group$Average.Package<-units_sold_by_week_by_contract_group$Units/units_sold_by_week_by_contract_group$no.of.Customers


year=2016
units_sold_by_week_by_contract_group<-numerical.description(units_sold_by_week_by_contract_group)


ggplot(data = subset(units_sold_by_week_by_contract_group, Year==year & Week.Sold>21 & Week.Sold<36 ), aes(x = Week.Sold, y = Average.Weekly.Units, group = Description)) +
    xlab("Week number") + ylab("Average Weekly Units") +
    geom_line(aes(color = Description),size=1.5)+scale_colour_brewer( palette="Paired")



#plotting
par(mfrow = c(2,2))
for (year in c(2014,2015,2016,2017)){
  ggplot(data = subset(units_sold_by_week_by_contract_group, Year==year & Week.Sold>21 & Week.Sold<36 ), aes(x = Week.Sold, y = Average.Weekly.Units, group = Description)) +
    xlab("Week number") + ylab(paste("Average Weekly Units",year)) +
    geom_line(aes(color = Description),size=1.5)+scale_colour_brewer( palette="Paired")
  
}


```
## Including Plots








```{r}
from_date="2014/01/01"
to_date="2014/12/31"

# just subset by year
#yearfourteen<-subset(data_frame, format(as.Date(Date.Sold),"%Y")==2014)

#How many sales were made by each contract group within given timeframe?
summary(data_frame[ (data_frame$Date.Sold >= from_date & data_frame$Date.Sold <= to_date )  , ]$Description)

#extract  all sales made in given timeframe.
customers_in_timeframe<-data_frame[ (data_frame$Date.Sold >= from_date & data_frame$Date.Sold <= to_date )  , ]

# 116xxxx  First Year
dealer_description_table=unique(data.frame(customers_in_timeframe$Dealer.Account.Number,customers_in_timeframe$Description))

#How many customers each dealer had
#table(customers_in_timeframe$Dealer.Account.Number)

# how many sales were made by contract type?
sales_by_contract<-table(customers_in_timeframe$Description)


#how many dealers did we have by contract year
dealers_by_contract<-table(dealer_description_table$customers_in_timeframe.Description)


library(gridExtra)
par(mfrow= c(1,2))
barplot(sales_by_contract,ylab ="Sales by contract", las=2)
barplot(dealers_by_contract,ylab = "Dealers by contract",las=2 )

#png(file="sales_by_contract.png",width=1024, height=600)

#barplot(sales_by_contract,ylab ="Sales by contract",las = 2)
dev.off()


grid.arrange(grobTree(sales_by_contract_plot),grobTree(dealers_by_contract_plot) , nrow=2)

```
You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
